---
title: "Diversity recovery"
author: "Nic McPhee and Tom Helmuth"
date: "23 March 2016"
output: html_document
---

# Setting things up

Let's load up the data:

```{r}
library("ggplot2")

div15run0 <- read.csv("data/RSWN/div15run0.csv")
div15run1 <- read.csv("data/RSWN/div15run1.csv")
div90run6 <- read.csv("data/RSWN/div90run6.csv")
div90run10 <- read.csv("data/RSWN/div90run10.csv")
drop25run0 <- read.csv("data/RSWN/drop25run0.csv")
drop25run8 <- read.csv("data/RSWN/drop25run8.csv")

dl_div15run0 <- read.csv("data/DoubleLetters/double_letters_div15_run0.csv")
dl_div15run1 <- read.csv("data/DoubleLetters/double_letters_div15_run1.csv")
dl_div90run0 <- read.csv("data/DoubleLetters/double_letters_div90_run0.csv")
dl_div90run3 <- read.csv("data/DoubleLetters/double_letters_div90_run3.csv")
dl_drop25run0 <- read.csv("data/DoubleLetters/double_letters_drop_25_run0.csv")
dl_drop25run21 <- read.csv("data/DoubleLetters/double_letters_drop_25_run21.csv")
```

Now create a new `selection` column to make it easier to organize the graphs. The code below is stupid ugly and depressingly repetitious. I tried writing a loop to factor out that repetition but it didn't work (the new rows got attached to the local variable instead of the "real" data frame), so I wimped out and copy/pasted like made. Sorry.

```{r}
#for (d in list(div15run0, div15run1, div90run6, div90run10, drop25run0, drop25run8)) {
#  d$selection <- sapply(strsplit(as.character(d$method), "-"), function(l) l[[3]])
#  print(names(d))
#}

div15run0$selection <- sapply(strsplit(as.character(div15run0$method), "-"), function(l) l[[3]])
div15run1$selection <- sapply(strsplit(as.character(div15run1$method), "-"), function(l) l[[3]])
div90run6$selection <- sapply(strsplit(as.character(div90run6$method), "-"), function(l) l[[3]])
div90run10$selection <- sapply(strsplit(as.character(div90run10$method), "-"), function(l) l[[3]])
drop25run0$selection <- sapply(strsplit(as.character(drop25run0$method), "-"), function(l) l[[3]])
drop25run8$selection <- sapply(strsplit(as.character(drop25run8$method), "-"), function(l) l[[3]])

dl_div15run0$selection <- sapply(strsplit(as.character(dl_div15run0$method), "-"), function(l) l[[3]])
dl_div15run1$selection <- sapply(strsplit(as.character(dl_div15run1$method), "-"), function(l) l[[3]])
dl_div90run0$selection <- sapply(strsplit(as.character(dl_div90run0$method), "-"), function(l) l[[3]])
dl_div90run3$selection <- sapply(strsplit(as.character(dl_div90run3$method), "-"), function(l) l[[3]])
dl_drop25run0$selection <- sapply(strsplit(as.character(dl_drop25run0$method), "-"), function(l) l[[3]])
dl_drop25run21$selection <- sapply(strsplit(as.character(dl_drop25run21$method), "-"), function(l) l[[3]])

div15run0$run <- sapply(strsplit(as.character(div15run0$method), "-"), function(l) l[[2]])
div15run1$run <- sapply(strsplit(as.character(div15run1$method), "-"), function(l) l[[2]])
div90run6$run <- sapply(strsplit(as.character(div90run6$method), "-"), function(l) l[[2]])
div90run10$run <- sapply(strsplit(as.character(div90run10$method), "-"), function(l) l[[2]])
drop25run0$run <- sapply(strsplit(as.character(drop25run0$method), "-"), function(l) l[[2]])
drop25run8$run <- sapply(strsplit(as.character(drop25run8$method), "-"), function(l) l[[2]])

dl_div15run0$run <- sapply(strsplit(as.character(dl_div15run0$method), "-"), function(l) l[[2]])
dl_div15run1$run <- sapply(strsplit(as.character(dl_div15run1$method), "-"), function(l) l[[2]])
dl_div90run0$run <- sapply(strsplit(as.character(dl_div90run0$method), "-"), function(l) l[[2]])
dl_div90run3$run <- sapply(strsplit(as.character(dl_div90run3$method), "-"), function(l) l[[2]])
dl_drop25run0$run <- sapply(strsplit(as.character(dl_drop25run0$method), "-"), function(l) l[[2]])
dl_drop25run21$run <- sapply(strsplit(as.character(dl_drop25run21$method), "-"), function(l) l[[2]])

div15run0$start <- sapply(strsplit(as.character(div15run0$method), "-"), function(l) l[[1]])
div15run1$start <- sapply(strsplit(as.character(div15run1$method), "-"), function(l) l[[1]])
div90run6$start <- sapply(strsplit(as.character(div90run6$method), "-"), function(l) l[[1]])
div90run10$start <- sapply(strsplit(as.character(div90run10$method), "-"), function(l) l[[1]])
drop25run0$start <- sapply(strsplit(as.character(drop25run0$method), "-"), function(l) l[[1]])
drop25run8$start <- sapply(strsplit(as.character(drop25run8$method), "-"), function(l) l[[1]])

dl_div15run0$start <- sapply(strsplit(as.character(dl_div15run0$method), "-"), function(l) l[[1]])
dl_div15run1$start <- sapply(strsplit(as.character(dl_div15run1$method), "-"), function(l) l[[1]])
dl_div90run0$start <- sapply(strsplit(as.character(dl_div90run0$method), "-"), function(l) l[[1]])
dl_div90run3$start <- sapply(strsplit(as.character(dl_div90run3$method), "-"), function(l) l[[1]])
dl_drop25run0$start <- sapply(strsplit(as.character(dl_drop25run0$method), "-"), function(l) l[[1]])
dl_drop25run21$start <- sapply(strsplit(as.character(dl_drop25run21$method), "-"), function(l) l[[1]])
```

# Plotting

This little guy sets changes the default color of outlier dots from black to whatever color is being used fo rthe boxplot:

```{r}
update_geom_defaults("point", list(colour = NULL, alpha=0.05))
```

## Replace-space-with-newline

These are runs starting after a dive in diversity in a lexicase selection run, presumably due to a hyperselection event.

```{r, echo=FALSE}
div15 = rbind(div15run0, div15run1)

ggplot(div15, aes(x=generation, y=error_vector_diversity, color=selection)) + 
  geom_boxplot(aes(group=interaction(generation, selection))) + 
  labs(title = "RSWN after diversity crash", x="Generation", y="Error diversity", color="Selection")
```

Now starting with a point in a lexicase selection run where the diversity was nice and high.

```{r, echo=FALSE}
div90 = rbind(div90run6, div90run10)

ggplot(div90, aes(x=generation, y=error_vector_diversity, color=selection)) + 
  geom_boxplot(aes(group=interaction(generation, selection))) + 
  labs(title = "RSWN starting with high diversity", x="Generation", y="Error diversity", color="Selection")
```

Lastly, starting with a point in a tournament selection run where diversity was fairly low.

```{r, echo=FALSE}
drop25 = rbind(drop25run0, drop25run8)

ggplot(drop25, aes(x=generation, y=error_vector_diversity, color=selection)) + 
  geom_boxplot(aes(group=interaction(generation, selection))) + 
  labs(title = "RSWN starting with low diversity", x="Generation", y="Error diversity", color="Selection")
```

## Double letters

These are runs starting after a dive in diversity in a lexicase selection run, presumably due to a hyperselection event.

```{r, echo=FALSE}
dl_div15 = rbind(dl_div15run0, dl_div15run1)

ggplot(dl_div15, aes(x=generation, y=error_vector_diversity, color=selection)) + 
  geom_boxplot(aes(group=interaction(generation, selection))) + 
  labs(title = "double-letters after diversity crash", x="Generation", y="Error diversity", color="Selection")
```

Now starting with a point in a lexicase selection run where the diversity was nice and high.

```{r, echo=FALSE}
dl_div90 = rbind(dl_div90run0, dl_div90run3)

ggplot(div90, aes(x=generation, y=error_vector_diversity, color=selection)) + 
  geom_boxplot(aes(group=interaction(generation, selection))) + 
  labs(title = "double-letters starting with high diversity", x="Generation", y="Error diversity", color="Selection")
```

Lastly, starting with a point in a tournament selection run where diversity was fairly low.

```{r, echo=FALSE}
dl_drop25 = rbind(dl_drop25run0, dl_drop25run21)

ggplot(dl_drop25, aes(x=generation, y=error_vector_diversity, color=selection)) + 
  geom_boxplot(aes(group=interaction(generation, selection))) + 
  labs(title = "double-letters starting with low diversity", x="Generation", y="Error diversity", color="Selection")
## Adding confidence intervals

This uses `ggplot2`'s `geom_smooth()` to create the 95% confidence interval _on the fitted curve_ (i.e., the curve is highly likely to be between the upper and lower bounds of the ribbon). This says nothing about where the actually data points will lie, however, which is I'm guessing is what Tom really wanted in a confidence interval.

```{r, warning=FALSE, message=FALSE}
ggplot(dl_drop25, aes(x=generation, y=error_vector_diversity, color=selection)) +
  geom_jitter(alpha=0.05) +
  geom_smooth(span=0.1) +
  facet_grid(. ~ run) +
  ylim(0, 1) +
  labs(title = "double-letters run 0 after diversity crash", x="Generation", y="Error diversity", color="Selection")
```
